---
title: "Estimation by Hospital"
author: "Zhuoheng Han, Ziyan Xia"
date: "4/22/2022"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

# Top-down Method Using PCA

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

We loaded the cleaned data and defined the total population of DRE patients based on the zip code dataset. 

```{r}
library(dplyr)
dat <- read.csv("~/Desktop/CMU/36-672_Statistical_Practice/LivaNova/Data_processing/cleaned_df_zip.csv")
#total_epi <- 3423550+150000*7
#total_DRE <- total_epi/3

total_DRE <- 1523075
```

According to the correlation plot, we found that most numerical variables are highly correlated

```{r}
library(corrplot)
dat_num <- dat %>% select(-c("ccn","cec","zip",'impute_discharge', 'impute_drg100','impute_drg101','impute_neuro1','impute_neuro2'))
corrplot(cor(dat_num), method = "number")
```


### Estimate the number of cec and non-cec DRE patients

We estimate the number of cec and non-cec patients using the sum of total discharges
```{r}
dat_0<-dat[which(dat$cec==0),]
dat_1<-dat[which(dat$cec==1),]
sum_0<-sum(dat_0$num_neuro_afl_1)
sum_1<-sum(dat_1$num_neuro_afl_1)
n_0<-(sum_0/(sum_0+sum_1))*total_DRE
n_1<-(sum_1/(sum_0+sum_1))*total_DRE
data.frame(non_cec_patients=n_0,cec_patients=n_1)%>% knitr::kable()
```

### Min-Max Normalization

The formula for a min-max normalization is$$\frac{X – min(X)}{max(X) – min(X)}$$

For each value of a variable, we simply find how far that value is from the minimum value, then divide by the range. 
We normalize all the numerical variables and use the average value of these normalized variables as our new variable. Using $$weight=\frac{this \ variable}{the\ sum\ of\ this\ variable}$$ we will get the population weight for each hospital and therefore could calculate the estimation of DRE patients for each hospital.

```{r}
min_max_norm <- function(x) {
    (x - min(x)) / (max(x) - min(x))
}
dat_0_norm <- as.data.frame(lapply(dat_0 %>% select(-c("ccn","cec","zip",'impute_discharge', 'impute_drg100','impute_drg101','impute_neuro1','impute_neuro2')), min_max_norm))
dat_1_norm <- as.data.frame(lapply(dat_1 %>% select(-c("ccn","cec","zip",'impute_discharge', 'impute_drg100','impute_drg101','impute_neuro1','impute_neuro2')) ,min_max_norm))

dat_0_norm$avg<-(dat_0_norm$total_discharges+dat_0_norm$drg_100+dat_0_norm$drg_101+dat_0_norm$total_beds_min+dat_0_norm$total_beds_max+dat_0_norm$num_neuro_afl_1+dat_0_norm$num_neuro_afl_2)/7

dat_0_norm$weights<-dat_0_norm$avg/sum(dat_0_norm$avg)
dat_0$estimation<-round(dat_0_norm$weights*n_0)

dat_1_norm$avg<-(dat_1_norm$total_discharges+dat_1_norm$drg_100+dat_1_norm$drg_101+dat_1_norm$total_beds_min+dat_1_norm$total_beds_max+dat_1_norm$num_neuro_afl_1+dat_1_norm$num_neuro_afl_2)/7

dat_1_norm$weights<-dat_1_norm$avg/sum(dat_1_norm$avg)
dat_1$estimation<-round(dat_1_norm$weights*n_1)

rbind(head(dat_0,5),head(dat_1,5))%>% knitr::kable()
```

```{r}
estimate_df <- rbind(dat_0,dat_1)
```

```{r}
location <- read.csv("~/Desktop/CMU/36-672_Statistical_Practice/LivaNova/Estimation_by_zipcode/uszips.csv")
all <- left_join(estimate_df, location[,1:3], by="zip")
all$cec<-as.factor(all$cec)
```

```{r}
library(leaflet)
leaflet(all) %>%
  addTiles() %>% addCircleMarkers(
    lat = ~lat, lng = ~lng,
    radius = sqrt(all$estimation)*.3,
    fillColor = ~ifelse(all$cec == 0,"#CD6155", "green"),
    stroke = FALSE, 
    fillOpacity = .3)
```

```{r}
pal <- colorNumeric(
  palette = 'RdYlBu',
  domain = all$estimation,
  n = 10
)

leaflet(all) %>%
  addTiles() %>% addCircleMarkers(
    lat = ~lat, lng = ~lng,
    radius = 3,
    fillColor = ~pal(estimation),
    stroke = FALSE) %>%
  addLegend("bottomright", pal = pal, values = ~estimation)
```


